### 2022-07-28 paper/localpsf.tex
some paper edits to shorten appendix. Almost sub 25 pages again

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.85]{mean_displacement_invariance.pdf}
	\end{center}
	\caption{$\Aop$ is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$. Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
	\label{fig:mean_displacement_invariance}
\end{figure}

@@

The action of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. We derive the extension as follows. Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
\begin{subequations}
	\label{eq:extension_to_distributions}
	\begin{align}
		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy 
		= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
	\end{align}
\end{subequations}
where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and instead suppose $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution. In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid. However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined. Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}. For convenience of notation, we denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.

@@

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
\begin{equation*}
	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
\end{equation*}
where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
	\label{eq:Ks_generalized}
	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.

To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's minimization algorithm \cite{Brent71} (any fast 1 dimensional convex optimization routine may be used). If $K(s^*) <0$ at the minimizer $s^*$, then $E_1 \cap E_2 = \{\}$. Otherwise $E_1 \cap E_2 \neq \{\}$.
%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

@@

One may reduce the average cost of performing ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect. If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary. If the bounding boxes intersect, the ellipsoids might intersect, or they might not intersect, so one must perform the complete ellipsoid intersection test described in Section \ref{sec:fast_ellipsoid_intersection_test}.

The smallest coordinate axis aligned box that contains the ellipsoid 
$
E = \{x : (x - \spatialmean)^T \spatialcov^{-1} (x - \spatialmean) \le \tau^2\}
$
is given as follows:
\begin{equation*}
	\prod_{i=1}^\gdim \left[ \spatialmean_i -\tau\sqrt{\Sigma_{ii}}, \spatialmean_i +\tau\sqrt{\Sigma_{ii}} \right],
\end{equation*}
where $\Sigma_{ii}$ is the $i$th diagonal entry of $\Sigma$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
$\prod_{i=1}^\gdim \left[a_i^-, a_i^+\right]$ and $\prod_{i=1}^\gdim \left[b_i^-, b_i^+\right]$
intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$.

@@

If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows
$
\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
$
If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite[Section 13.8]{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. In detail, let $\Pi$ be the indicator function for the interval $(a,b)$,
\begin{equation*}
	\Pi(\lambda) = \begin{cases}
		1, & \lambda \in (a,b), \\
		0, & \text{otherwise}.
	\end{cases}
\end{equation*}
Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
$
\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
$
The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$. If we choose $a < \lambda_\text{min}$ and $b=0$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues, and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero. Subtracting a negative number from itself twice flips its sign, so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
\begin{equation}
\label{eq:I_Pi_abs}
	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
\end{equation}
where $\mathbf{I}$ is the identity matrix of the appropriate size.
By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.

>>>>

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.8]{mean_displacement_invariance.pdf}
	\end{center}
	\caption{$\Aop$ is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$. Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
	\label{fig:mean_displacement_invariance}
\end{figure}

@@

The action of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. 
%We derive the extension as follows. 
Let $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
\begin{subequations}
	\label{eq:extension_to_distributions}
	\begin{align}
		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy 
		= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
	\end{align}
\end{subequations}
where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution. In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid. However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined. Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}. For convenience of notation, we denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.

@@

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
$
	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
$
where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
	\label{eq:Ks_generalized}
	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$. To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's minimization algorithm \cite{Brent71}. If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$. Otherwise $E_1 \cap E_2 \neq \{\}$.
%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

@@

One may reduce the average cost of performing ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect. If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary. If the bounding boxes intersect, the ellipsoids might intersect, or they might not intersect, so one must perform the complete ellipsoid intersection test described in Section \ref{sec:fast_ellipsoid_intersection_test}. The smallest coordinate axis aligned box that contains the ellipsoid 
$
E = \{x : (x - \spatialmean)^T \spatialcov^{-1} (x - \spatialmean) \le \tau^2\}
$
is the box
$
	\prod_{i=1}^\gdim \left[ \spatialmean_i -\tau\sqrt{\Sigma_{ii}}, \spatialmean_i +\tau\sqrt{\Sigma_{ii}} \right],
$
where $\Sigma_{ii}$ is the $i$th diagonal entry of $\Sigma$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
$\prod_{i=1}^\gdim \left[a_i^-, a_i^+\right]$ and $\prod_{i=1}^\gdim \left[b_i^-, b_i^+\right]$
intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$.

@@

If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows
$
\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
$
If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite[Section 13.8]{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. In detail, let $\Pi$ be the indicator function for the interval $(a,b)$. That is, $\Pi(\lambda)=0$ for $\lambda \in (0,1)$, and $\Pi(\lambda)=0$ otherwise.
%\begin{equation*}
%	\Pi(\lambda) = \begin{cases}
%		1, & \lambda \in (a,b), \\
%		0, & \text{otherwise}.
%	\end{cases}
%\end{equation*}
Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
$
\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
$
The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$. If we choose $a < \lambda_\text{min}$ and $b=0$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues, and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero. Subtracting a negative number from itself twice flips its sign, so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
\begin{equation}
\label{eq:I_Pi_abs}
	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
\end{equation}
where $\mathbf{I}$ is the identity matrix of the appropriate size.
By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
